### 2015-09-12 vmcai-2016-tex/bisimulation.tex
All sections were edited to remove first-person voice, e.g. we, our, etc.
Revision and refactor

\section{Correctness}
\label{sec:bisim}
A major drawback of previous methods for creating test inputs for
symbolic heaps has been a lack of completeness. Incomplete heap
representations may not create a set of inputs that represents
the entire space of possible inputs. By missing some possible
inputs, the subsequent analysis may fail to recognize 
possible program behaviors.

The theorems in this section establish the soundness and
completeness of the symbolic heap approach with respect to Generalized
Symbolic Execution (\gsetxt{}). Intuitively, the theorems imply that any properties proven
with~\gsetxt{} can also be proved using the symbolic heap
algorithm. The proofs for the theorems, as well as
the complete set of semantic rules for~\gsetxt{} and~\symtxt{} can
be found in ~\cite{Hillery:2015}.

The proof shows the existence of a
bisimulation between sets of states related by~\gsetxt{} ($p \rgse
p^\prime$) and states related by the symbolic heap and update rules in this paper ($q \rsym
q^\prime$). The relations are on the universe of \emph{well-formed}
states $S$, which have the properties in \secref{sec:sh} with the
constraint that the states are \emph{feasible}: successors exist
unless at \cfgt{end}. Let $p \rgse p^\prime$ be a union over relations
for~\gsetxt{} (see~\cite{Hillery:2015}): $\rgse = \rgse^\mathit{A} \cup
\rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup
\rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup
\rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access
after evaluating the expression for the base reference $\lp
\cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp$, \emph{W} is
a field write after evaluating the expression for the right operand
$\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Leftrightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose we have a set $P$ containing a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in~\gsetxt{}: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.

>>>>

\section{Correctness}
\label{sec:bisim}

Incomplete heap representations may not create a set of inputs that represents
the entire space of possible inputs. By missing some possible
inputs, the subsequent analysis may fail to recognize 
possible program behaviors. The theorems in this section establish the soundness and
completeness of symbolic initialization with respect to classical GSE with lazy initialization. Intuitively, the theorems imply that any properties proven
with~\gsetxt{} can also be proved using symbolic initialization. The proofs for the theorems, as well as
the complete set of semantic rules for~\gsetxt{} and~\symtxt{} can
be found in ~\cite{Hillery:2015}.

The proof shows the existence of a
bisimulation between sets of states related by~\gsetxt{} ($p \rgse
p^\prime$) and states related by the symbolic heap and update rules in this paper ($q \rsym
q^\prime$). The relations are on the universe of \emph{well-formed}
states $S$, which have the properties in \secref{sec:sh} with the
constraint that the states are \emph{feasible}: successors exist
unless at \cfgt{end}. Let $p \rgse p^\prime$ be a union over relations
for~\gsetxt{} (see~\cite{Hillery:2015}): $\rgse = \rgse^\mathit{A} \cup
\rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup
\rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup
\rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access
after evaluating the expression for the base reference $\lp
\cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp$, \emph{W} is
a field write after evaluating the expression for the right operand
$\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Leftrightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose a set $P$ contains a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in~\gsetxt{}: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.

### 2015-09-12 vmcai-2016-tex/intro.tex
Minor tweak to the intro.
Revision and refactor

A common hurdle to using guarded value sets is the treatment of symbolic heap inputs. For many applications, it is desirable to model the input heap as a black box, with operations over the input references providing the only means of divining its structure. Precisely modeling operations on black-box heap inputs is challenging because it requires formulating logical predicates over an input domain that contains a potentially unbounded number of hidden references~\cite{Chen:2013,Qu:2011}. Generalized symbolic execution (\gsetxt{}) methods~\cite{GSE03,Deng:2007} provide an accurate solution, but they quickly produce an overwhelming number of execution paths for all but the simplest heap shapes. Efforts to extend~\gsetxt{} principles to guarded value sets have failed to produce provably equivalent results. Thus, despite the advantages of using guarded value sets, it remains an open question how they can be used in symbolic execution to automatically model all possible program behaviors in the case of an arbitrary input heap.

The contribution of this work is to build on previous efforts, by demonstrating an initialization scheme that enables pointer dereferencing in an unknown symbolic input heap, and by proving that the initialization scheme is sound and complete with respect to properties provable by symbolic execution. A proof-of-concept implementation also demonstrates the efficiency of value-summary representations, as compared to existing case-splitting methods.

>>>>

A common hurdle to using guarded value sets is the treatment of symbolic heap inputs. For many applications, it is desirable to model the input heap as unconstrained, with operations over the input references providing the only means of divining its structure. Precisely modeling operations on unconstrained heap inputs is challenging because it requires formulating logical predicates over an input domain that contains a potentially unbounded number of hidden references~\cite{Chen:2013,Qu:2011}. Generalized symbolic execution (\gsetxt{}) methods~\cite{GSE03,Deng:2007} provide an accurate solution, but they quickly produce an overwhelming number of execution paths for all but the simplest heap shapes. Efforts to extend~\gsetxt{} principles to guarded value sets have failed to produce provably equivalent results \cite{LISTEFFORTS}. Thus, despite the advantages of using guarded value sets, it remains an open question how they can be used in symbolic execution to automatically model all possible program behaviors in the case of an arbitrary input heap.

The contribution of this work is a symbolic initialization scheme for uninitialized pointers in a fully symbolic heap. Where GSE lazily instantiates uninitialized pointers to either reference NULL, a new instance of the correct type, or an alias to a previously initialized object, the symbolic initialization in this work creates a guarded value set expressing all of these eventualities in the same symbolic heap. Where GSE branches the search space on each choice, symbolic initialization does not. Symbolic initialization only branches on reference compares, partitioning the symbolic to indicate heaps where the references alias and heaps where the references do not.

This paper includes a proof that symbolic initialization is sound and complete with respect to properties provable by GSE. Such a results is important because it means that symbolic initialization can be used to create input heaps for other analyses in a way that is provably correct with regards to GSE: there are no missing heaps and there are no extra heaps.

A proof-of-concept implementation of symbolic initialization is reported in this paper. It demonstrates the efficiency of~\symtxt{}, as compared to existing~\gsetxt{} methods. The use of the guarded value sets does require extra interaction with an SMT solver, but the savings in representing multiple heap shapes in a single representation overcomes the cost of the solver calls. In fact, the comparison shows an exponential increase in the number of heaps that can be analyzed in a set of benchmark examples meaning that other approaches based on~\gsetxt\ fail to complete.

### 2015-09-12 vmcai-2016-tex/lazy.tex
Revision and refactor

\subsection{Generalized Symbolic Execution}
\label{app:gse}
The rules to generalized symbolic execution with
lazy initialization using the symbolic heap are in
\figref{fig:lazyInit} for the initialization and \figref{fig:lazy} for
how the initialization is used. Specifically, the field-access rule in
\figref{fig:lazy} uses initialization, $\rightarrow_I^*$, to ensure
the accessed field is instantiated. Initialization in generalized
symbolic execution never happens for more than one object on any use
of $\rightarrow_I^*$: $\Lambda$ is either empty or contains exactly
one location.

The~\gsetxt{} initialization rules preserve several important properties of the symbolic heap. In the heap generated by~\gsetst{}, each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals}.
  Only input heap references appear in constraints to express potential aliasing.
  Auxiliary literals and stack references are used to define additional structure in the heap but only stack references can appear in environments, expressions, or continuations. Changes to $R$ are monotonically increasing, eliminating any need for constraint rewriting as the program state evolves.
A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

An uninitialized object may be lazily initialized
non-deterministically to null, a new instance of the object, or an to
an object of the same type also initialized lazily. The
non-deterministic initialization creates branches in the execution
tree during state space exploration.

The $\rgse$ relation is a union over branching points on control flow
to support the final proof of the existence of a
bisimulation. Although it is apparent that there is non-determinism in
initialization, it is not the initialization where control flow
diverges in program execution. Rather, that occurs at branches and
exceptions, and that is reflected in the composition of $\rgse$.

>>>>

\section{GSE}
\label{app:gse}
Classical GSE and its variants all subscribe to the same ideal: an uninitialized object may be lazily initialized
in a non-deterministic way to null, a new instance of the object, or an to
an object of the same type also initialized lazily (i.e., the \emph{input heap}) \cite{GSE03,KiasanKunit,Cadar:2008,Rosner:2015}. The
non-determinism creates branches in the execution
tree during state space exploration. GSE semantics are defined here to give insight to symbolic initialization and support the proof that it preserves GSE behaviors.

The rules for classical GSE using the symbolic heap are in
\figref{fig:lazyInit} for lazy initialization: NULL, new, and alias. Fresh references and locations strictly increase so it is possible to minimize over a set to find the first created. References are partitioned to support latter proofs: $\mathrm{init}_\cfgnt{r}()$ for the input
  heap
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
    literals}.
  In general, as shown in the next section,
  only input heap references appear in constraints to express potential aliasing, and only stack references can appear in environments, expressions, or continuations.  Finally, the \emph{isInit} is true for initialized references from the input heap (i.e., potential aliases).

How the lazy
initialization is used is defined in \figref{fig:lazy}. Specifically,
the field-access rule in \figref{fig:lazy} uses initialization,
$\rightarrow_I^*$, to ensure the accessed field is
instantiated. Initialization in generalized symbolic execution never
happens for more than one object on any use of $\rightarrow_I^*$:
$\Lambda$ is either empty or contains exactly one location. The rest of the rules in \figref{fig:lazy} do not initialize, but they are included to elucidate how symbolic initialization differs from GSE with lazy initialization.

The $\rgse$ relation defined in \figref{fig:lazy} is a union over branching points on control flow
to support the final proof of the existence of a
bisimulation. Although it is apparent that there is non-determinism in
initialization, it is not the initialization where control flow
diverges in program execution. Rather, that occurs at branches and NULL-exceptions, and that is reflected in the composition of $\rgse$.

### 2015-09-12 vmcai-2016-tex/summary-heap.tex
Revision and refactor

\section{Prelimiaries}
\label{sec:sh}

\subsection{Symbolic Execution}

Symbolic execution semantics are defined using
a small-step operational semantics similar to the one presented 
in~\cite{saints-MS}. The small-step semantics,
defined as a syntactic machine using a CESK architecture~\cite{Felleisen:1992},
are modified to perform symbolic execution on a symbolic heap. The
surface (input) syntax and machine (state) syntax are shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.

@@

A \emph{symbolic heap} is a tuple $(\cfgnt{L}\ \cfgnt{R}\ \phi\ \eta)$, where the~\emph{location map} $\cfgnt{L}$,~\emph{reference map} $\cfgnt{R}$, and~\emph{environment} $\eta$ are partial functions, and the~\emph{path condition} $\phi$ is a predicate over the program variables. Given reference $r$, $\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the guarded value set in the heap associated with $r$. The constraint  associated with each location is a guard
representing the conditions under which the associated reference
maps to the given location. By definition, any location appears 
at most once in any given value set. Unions of value sets containing
the same location are resolved by forming a disjunction of the constraints
from each set. $\cfgnt{R}(\cfgnt{l}, \cfgnt{f}) = \cfgnt{r}$ is the
reference associated with the given location-field pair in the
heap. The environment  $\eta$ maps local variables to references.

Conceptually, the symbolic heap may be thought of as a bipartite graph. \figref{fig:exampleHeap} shows an example symbolic heap graph. References are 
represented by circles and locations are represented by squares. Arrows leaving
references correspond to the guarded value sets returned by the $\cfgnt{L}$ function, and 
arrows leaving the squares correspond to the $\cfgnt{R}$ function.

The machine state $\cfgnt{s}$ includes the program
registry $\mu$, the symbolic heap,
and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation.  The
more complex structures, such as the environment and heap, are defined
as singly linked lists of mappings that start with empty, \cfgnt{mt}. These lists are treated as
partial functions so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. The continuation
$\cfgnt{k}$ indicates with the symbol $\cfgnt{*}$ where the expression $\cfgnt{e}$ came from, stores
temporary computation, and keeps track of the next continuation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.

Execution begins with the initial machine state. The expression for the initial state constructs a new object of type
$\cfgnt{C}$ and then invokes the method $\cfgnt{m}$ in that
object. The initial symbolic heap contains 
two special locations, null ($\cfgnt{l}_\mathit{null}$), and uninitialized
  ($\cfgnt{l}_\mathit{un}$),  with corresponding references 
  $\cfgnt{r}_\mathit{null}$ and $\cfgnt{r}_\mathit{un}$ 
  where $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
Given an input program $\lp\mu\ \lp\cfgnt{C}\ \cfgnt{m}\rp\rp$, the expression for the initial machine state is
$$
\begin{array}{l}
\lp\mu 
\ \cfgnt{L}[\cfgnt{r}_\mathit{null} \mapsto \{\lp\cfgt{true}\ \cfgnt{l}_\mathit{null})\}\rp] 
           [\cfgnt{r}_\mathit{un} \mapsto \{\lp\cfgt{true}\ \cfgnt{l}_\mathit{un}\rp\}] \\
\ \cfgnt{R}
\ \cfgt{true}\ \eta\  \lp\lp\cfgt{new}\ \cfgnt{C}\rp\ \cfgt{@}\ \cfgnt{m}\ \cfgt{true}\rp\rp\ \cfgt{end}\rp
\end{array}
$$

The semantics are expressed as
rewrites on strings using pattern matching. 
Consider the rewrite rule
for the beginning of a field access instruction:
$$
\mprset{flushleft}
	\inferrule[Field Access(eval)]{}{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgnt{e}\ \cfgt{\$}\ \cfgnt{f}\rp \ \cfgnt{k}\rp  \rcom 
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \cfgnt{e}\ \lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp \rp 
	}
$$
If the string representing the current state matches the left side, then it
creates the new string on the right. In this example, the new string
on the right is now evaluating the expression $\cfgnt{e}$ in the field
access, and it includes the continuation indicating that it still
needs to complete the actual field access once the expression is
evaluated.
%
%Other more complex rules, such as one to create a new instance of a
%class, define constraints on the rewrites and more complex
%transformations on the heap.
%$$
%\mprset{flushleft}
%	\inferrule[New]{
%      \cfgnt{r} = \mathrm{stack}_r\lp \rp\\
%      l = \mathrm{fresh}_l\lp \cfgnt{C}\rp\\\\
%      \cfgnt{R}^\prime = \cfgnt{R}[\forall \cfgnt{f} \in \mathit{fields}\lp \mathrm{C}\rp \ \lp \lp l\ \cfgnt{f}\rp  \mapsto \cfgnt{r}_\mathit{null} \rp ] \\\\
%      \cfgnt{L}^\prime = \cfgnt{L}[\cfgnt{r} \mapsto \{\lp \cfgt{true}\ l\rp \}]
%    }{
%      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgt{new}\ \cfgnt{C}\rp \ \cfgnt{k}\rp  \rcom
%      \lp \cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \phi\ \eta\ \cfgnt{r}\ \cfgnt{k}\rp 
%	}
%$$ In this rule, when the string matches the new-expression, it is
%        rewritten to use a new heap location where all of the fields
%        for the new object point to $\cfgnt{r}_\mathit{null}$ and the
%        location map points a new stack reference to that new object.
%        New heap locations returned from $\mathrm{fresh}_l$ are monotonically
%        increasing.
%        Also note that references for objects from the new-expression are
%        auxiliary literals, so they do not appear in any constraint in
%        the symbolic heap as an aliasing option. 
%        
% \input{javalite-common}
%

The rewrite relations for the portion of the language
semantics that are common to both generalized symbolic execution and
the symbolic heap algorithm may be found in \cite{Hillery:2015}. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking. The
machine halts if no rewrite is enabled.

>>>>

\section{A Symbolic Heap}
A \emph{symbolic heap} is a tuple
$(\cfgnt{L}\ \cfgnt{R}\ \phi\ \eta)$, where the~\emph{location map}
$\cfgnt{L}$,~\emph{reference map} $\cfgnt{R}$, and~\emph{environment}
$\eta$ are partial functions, and the~\emph{path condition} $\phi$ is
a predicate over the program variables. Given reference $r$,
$\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the guarded value set
in the heap associated with $r$. The constraint associated with each
location is a guard representing the conditions under which the
associated reference maps to the given location. By definition, any
location appears at most once in any given value set. Unions of value
sets containing the same location are resolved by forming a
disjunction of the constraints from each set. $\cfgnt{R}(\cfgnt{l},
\cfgnt{f}) = \cfgnt{r}$ is the reference associated with the given
location-field pair in the heap.

Conceptually, the symbolic heap may be thought of as a bipartite
graph. \figref{fig:exampleHeap} shows an example symbolic heap
graph. References are represented by circles and locations are
represented by squares. Arrows leaving references correspond to the
guarded value sets returned by the $\cfgnt{L}$ function, and arrows
leaving the squares correspond to the $\cfgnt{R}$ function. The
reference $\cfgnt{r}_1$ has has two members in it guarded value set,
so the location pointed to by $\cfgnt{r}_1$ depends on its aliasing
relationship to $\cfgnt{r}_0$.

The reference $\cfgnt{r}_\mathit{un}$ is a special reference to
indicate something that has yet to be initialized. In general,every
symbolic heap contains two special locations, null
($\cfgnt{l}_\mathit{null}$), and uninitialized
($\cfgnt{l}_\mathit{un}$), with corresponding references
$\cfgnt{r}_\mathit{null}$ and $\cfgnt{r}_\mathit{un}$ where
$\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
\{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
$\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
\{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$.

For well-formed symbolic heaps, a reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. Also, for a well-formed heap, all locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

\section{Operational Semantics}
This paper defines symbolic initialization using
a small-step operational semantics in the context of a syntactic machine with a CESK architecture~\cite{Felleisen:1992, saints-MS}. The
surface (input) syntax and machine (state) syntax (which omits the symbolic heap definition previously given and the environment) are shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.

@@

The machine state $\cfgnt{s}$ includes the program
registry $\mu$, the symbolic heap, the global path constraint, the environment $\eta$, the current expression (i.e., program), and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation. The environment is a partial function so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. This notation is used with symbolic heaps as well.

The continuation
$\cfgnt{k}$ indicates with the symbol $\cfgnt{*}$ where the expression $\cfgnt{e}$ came from, stores
temporary computation, and keeps track of the next continuation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.

Given an input program $\lp\mu\ \lp\cfgnt{C}\ \cfgnt{m}\rp\rp$, the expression for the initial machine state is
$$
\begin{array}{l}
\lp\mu 
\ \cfgnt{L}[\cfgnt{r}_\mathit{null} \mapsto \{\lp\cfgt{true}\ \cfgnt{l}_\mathit{null})\}\rp] 
           [\cfgnt{r}_\mathit{un} \mapsto \{\lp\cfgt{true}\ \cfgnt{l}_\mathit{un}\rp\}] \\
\ \cfgnt{R}
\ \cfgt{true}\ \eta\  \lp\lp\cfgt{new}\ \cfgnt{C}\rp\ \cfgt{@}\ \cfgnt{m}\ \cfgt{true}\rp\rp\ \cfgt{end}\rp
\end{array}
$$
The semantics are expressed as
rewrites on strings using pattern matching. 
Consider the rewrite rule
for the beginning of a field access instruction:
$$
\mprset{flushleft}
	\inferrule[Field Access(eval)]{}{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgnt{e}\ \cfgt{\$}\ \cfgnt{f}\rp \ \cfgnt{k}\rp  \rcom 
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \cfgnt{e}\ \lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp \rp 
	}
$$
If the string representing the current state matches the left side, then it
creates the new string on the right. In this example, the new string
on the right is now evaluating the expression $\cfgnt{e}$ in the field
access, and it includes the continuation indicating that it still
needs to complete the actual field access once the expression is
evaluated.
%
%Other more complex rules, such as one to create a new instance of a
%class, define constraints on the rewrites and more complex
%transformations on the heap.
%$$
%\mprset{flushleft}
%	\inferrule[New]{
%      \cfgnt{r} = \mathrm{stack}_r\lp \rp\\
%      l = \mathrm{fresh}_l\lp \cfgnt{C}\rp\\\\
%      \cfgnt{R}^\prime = \cfgnt{R}[\forall \cfgnt{f} \in \mathit{fields}\lp \mathrm{C}\rp \ \lp \lp l\ \cfgnt{f}\rp  \mapsto \cfgnt{r}_\mathit{null} \rp ] \\\\
%      \cfgnt{L}^\prime = \cfgnt{L}[\cfgnt{r} \mapsto \{\lp \cfgt{true}\ l\rp \}]
%    }{
%      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgt{new}\ \cfgnt{C}\rp \ \cfgnt{k}\rp  \rcom
%      \lp \cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \phi\ \eta\ \cfgnt{r}\ \cfgnt{k}\rp 
%	}
%$$ In this rule, when the string matches the new-expression, it is
%        rewritten to use a new heap location where all of the fields
%        for the new object point to $\cfgnt{r}_\mathit{null}$ and the
%        location map points a new stack reference to that new object.
%        New heap locations returned from $\mathrm{fresh}_l$ are monotonically
%        increasing.
%        Also note that references for objects from the new-expression are
%        auxiliary literals, so they do not appear in any constraint in
%        the symbolic heap as an aliasing option. 
%        
% \input{javalite-common}
%

The rewrite relations for the more mundane portions of the language
that do not update the symbolic heap are in the appendix with more details in 
\cite{Hillery:2015}. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking. The
machine halts if no rewrite is enabled. In the rest of this paper the relation $s \rcom
s^\prime$ indicates that two states are related by these more
mundane rules.
